\documentclass[10pt]{article} 
\usepackage{geometry, hyperref, booktabs}
\geometry{tmargin=1in, bmargin=1in, lmargin=2.5in, rmargin = 1in}  

\title{\textbf{U.S. Congress}\\\textit{Discussing the Electoral Connection}}
\author{Nathan Barron}
\date{Fall 2024}

\begin{document}
\maketitle
\tableofcontents
\vspace{.25in}
\hrule
\vspace{.25in}


\section{Analyzing Congressional behavior}
\subsection{The proximate goal}
\begin{itemize}
    \item What is Mayhew's primary argument or assertion?
    \item Mayhew begins with three questions. What are they and what are his answers? 
\end{itemize}
\subsection{Against parties}
\begin{itemize}
    \item What is the Downsian model of parties?
    \item Mayhew describes three reasons why the British system demonstrates high party loyalty and the U.S. system does not. What are they?
    \item What is Mayhew's counterpoint to the expected party differential?
    \item What is Mayhew's discussion about marginal MCs? What does he argue? 
\end{itemize}
\subsection{The safety myth}
\begin{itemize}
    \item Are members in a position to do anything about getting reelected? 
    \item What are the three campaign resources that Mayhew mentions?
    \item According to Mayhew, do MCs seek to maximize their electoral gains? Why or why not?
    \item Mayhew points out that, despite the incumbency advantage, MCs act like they are unsafe. Why is this?
\end{itemize}
\subsection{Three tools of incumbency}
\begin{itemize}
    \item What are the three actions that Mayhew mentions? What are the differences between them.
    \item Advertising has little or no issue content. What are some examples? Why is it helpful? How effective do you think advertising is (and why)? 
    \item Credit claiming is taking responsibility for positive government actions. What are the two things that Mayhew mentioned that MCs might take credit for? Mayhew says that these claims should be credible. Do they? What makes a credit claim credible? 
    \item Position taking is registering one's position on an issue important to relevant political actors. What are the benefits of position taking? Risks? What does Mayhew suggest is the most viable strategy for most MCs?
    \item There are general trends in who engages in what types of activities. Mayhew mentions that Senators tend to engage in position taking while House members tend to focus on credit claiming. Recent research has also found that moderates tend to focus on credit claiming while the less moderate representatives tend to focus on position taking. What are some potential consequences of these different behavior patterns? 
\end{itemize}
\end{document}
